---
title: "Reading external survey data"
author: "Arnob L. ALam"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Load Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Purpose

This vignette shows how to read the data provided by 
[(Fafchamps and Comola, 2013)] [1] and save the data is a .RData file that can
bse used in later analysis.

# Setting up R

We first create some variables to store things like the path to the
root of the project and the external data folder.  The vairables are as follows

* `baseLoc` - The location of the "root" of the project (all files and folders
    related to the project are below this folder).
* `extPath` - The location of "external data".  Example of external data is
    the result from the Nyakatoke village survey saved in CSV format.
* `dataPath` - The location of "internal data".  This is data that has been
    taken from the external source and converted to R format. For example, the
    Nyakatoke village data (from the external data folder) saved in .RData
    format.

```{r setup}
baseLoc <- system.file(package="riskSharing")
extPath <- file.path(baseLoc, "extdata")
dataPath <- file.path("data")
```

# Reading and saving data

The Nyakatoke village survey data is saved in the file "tanzanaia_data.csv" in
the external data folder.  This data is obtained by converting the Stata file
provided by [(Fafchamps and Comola, 2013)] [1]

```{r read-and-save, eval=FALSE}
nyakatoke <- read.csv(file.path(extPath,"tanzania_data.csv"))
save(nyakatoke, file = file.path(dataPath, "nyakatoke.RData"))
```
[1]: http://onlinelibrary.wiley.com/doi/10.1111/ecoj.12071/abstract "Comola, M. and Fafchamps, M. (2014), Testing Unilateral and Bilateral Link Formation. Econ J, 124: 954–976. doi:10.1111/ecoj.12071"