---
title: "Basic Descriptive Statistics"
author: "Arnob L. Alam"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Descriptive Statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Purpose

This vigentte shows some basic descriptive statstics about the networks fomred
by the Nyakatoke Village survey.

# Survey and data organization

The survey data comes from (Fafchamps and Comola, 2013) [1].  They provide the
following description of the data

> The Tanzanian data set comes from a village community named Nyakatoke in the 
Buboka Rural District of Tanzania, at the west of Lake Victoria. The data set is 
a census covering all 119 households in the village and includes information on 
households' demographics, wealth and assets, income sources and income shocks, 
transfers and interpersonal relations...

> During the first survey round, each adult in the village was asked: 
‘Can you give a list of people from inside or outside of Nyakatoke, who you can 
personally rely on for help and/or that can rely on you for help in cash, kind 
or labour?’Aggregated at the level of each household, the responses to this 
question constitute variables $g_{ij}^{i}$ and $g_{ji}^j$. In other words,
$g_{ij}^{i} = 1$ if an adult member of household $i$ mentions an adult member 
of household $j$ in their response to the above question.

In the data, the households are given a number from 1 through 119.  The columns 
`hh1` and `hh2` are the IDs of two households. `willingness_link1` takes on a 
value of $1$ if at least one adult member from the household identified by `hh1` 
mentioned someone from the household identified by `hh2` in their answer to the 
question "Can you give a list of people from inside or outside of Nyakatoke, 
who you can  personally rely on for help and/or that  can rely on you for help 
in cash, kind or labour"?  Similarly, `willingness_link2` takes on a value of 1 
if a member of `hh2` listed someone from `hh1` to the same question.

```{r short description of data}
library(riskSharing)
data("nyakatoke")
knitr::kable(head(nyakatoke[,1:4]))
```

Table: The first few rows and four columns of data from the Nyakatoke village
    survey
    
Here we see (from row 1) that household 113 and household 1 have no link betwwen 
them. From row 2, we see Household 76 and Household 1 might have a link because 
someone from household 1 mentioned household 76 in the answer to the survey.
However, no one from household 76 mentioned anyone from household 1 in their
answer.  Finally, from row 3 we see there is a link between household 17 and
household 1 (both `willingness_link1` and `willingness_link2` are 1).
    
The definitions of these columns will be expanded upon when we use them in
later analysis.

Note that in our dataset, each pair of household shows up twice, once in the `hh1`
column and again in the `hh2` column.  For example, household 1 and household 76
show up as:

```{r households pairs appear twice}
knitr::kable(subset(nyakatoke[,1:4], (hh1 == 1 & hh2 == 76) | (hh1 == 76 & hh2 == 1)))
```

Here we see the combination (household 76, household 1) show up in row 2 of the data
and the combination (household 76, household 1) show up in row 8937 of the data.
Looking at the `willingness_link1` and `willingness_link2` columns, we see that the
data in these rows are just flipped versions of each other.

The data also lists other characteristics of the housholds.  Here we show a few
rows and all columns of the dataset.

```{r head all columns}
knitr::kable(head(nyakatoke))
```

Table: All columns from in the data.

# Several networks from underlying data

Using the data from the survey, we can create several networks.  We create the
following three networks (more details about what each network represents is
provided in our paper).

1. A _directed network_.  If (an adult member of) household 1 lists (somenoe from)
    household 2 as an answer to the relevant survey question, we say there is a 
    directed link FROM household 1 TO household 2. This corrsponds to the desire 
    to link model in our paper.

2. An _undirected network assuming links are underreported_.  In this case, if an
    (adult member) from household 1 mentions someone from household 2 as an 
    answer to the relevant survey question, we say there is an undirected link
    between the two households.  This corresponds to the underreporting model
    in our paper.

3. An _undirected network assuming links are overreported_. In this case, if 
    members from both households list someone from the other household as an
    answer to the relevant survey question, we say there is an undirected link
    between the two households.  This corresponds to the overreporting model
    in our paper.
    
```{r create networks}
# For the desire to link data, we create an "edgelist" that igraph can read.
# This is a list of households where the first household in the pair mentioned
# someone from household 2 in their survey response.
edgelist <- as.matrix(nyakatoke[nyakatoke$willingness_link1 == 1, 1:2])
g.directed <- igraph::graph_from_data_frame(edgelist)

# For the underreporting model, we create a graph from the dataframe
# First we keep only the rows where at least one houshold said the other is in
# their network.
underreporting.df <- 
    nyakatoke[nyakatoke$willingness_link1 == 1 | nyakatoke$willingness_link2 == 1,]
# We then convert this into an undirected graph
g.underreporting <- igraph::graph_from_data_frame(underreporting.df, 
                                                  directed = FALSE)
# We then remove any multiple edges
g.underreporting <- igraph::simplify(g.underreporting)

# For the overreporting model we only keep those rows where both households
# listed the other in their network
overreporting.df <- 
    nyakatoke[nyakatoke$willingness_link1 == 1 & nyakatoke$willingness_link2 == 1,]
# We then create an graph
g.overreporting <- igraph::graph_from_data_frame(overreporting.df, directed = FALSE)
# From examining the list of vertices, we note that households 7, 30 32, 36, 44,
# 65, 84, 88, 91, 96, 107, 110, 116, 117, 118 and 119 don't show up in the 
# graph. This is because there is no case where these households list another 
# household in their support network and they do the same.  We dd back these 
# vertices as isolated vertices
missing.vertices <- c(7, 30, 32, 46, 44, 65, 84, 88, 91, 96, 107, 110, 
                      116, 117, 118, 119)
g.overreporting <- (Reduce(f = function(x, y) {y + igraph::vertex(x)},
       x = missing.vertices,
       init = g.overreporting,
       right = TRUE))
# We remove multiple edges
g.overreporting <- igraph::simplify(g.overreporting)
```

Now that we have created the graphs, we report some basic descriptive statistics
for each graph:

1. The number of vertices
2. The number of edges
3. The average degree (or average in-degree and out-degree for a directed graph)
4. The average path length

## Directed Graph

```{r descriptive stats directed}
descriptive_stats(g.directed)
```

## Undirected graph

### Underreporting model

```{r descriptive stats under}
descriptive_stats(g.underreporting)
```

### Overreporting Model
```{r descriprive stats over}
descriptive_stats(g.overreporting)
```