---
title: "Graph simulations"
author: "Arnob Alam"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Purpose

In this vignette we simulate some random graphs and compare the results with
the observed network

# Preliminary

As a reminder, we recreate the graphs (from the basic descriptive statistics
vignette)

```{r}
library(riskSharing)
data("nyakatoke")

edgelist <- as.matrix(nyakatoke[nyakatoke$willingness_link1 == 1, 1:2])
g.directed <- igraph::graph_from_data_frame(edgelist)

underreporting.df <- 
    nyakatoke[nyakatoke$willingness_link1 == 1 | nyakatoke$willingness_link2 == 1,]
g.underreporting <- igraph::graph_from_data_frame(underreporting.df, 
                                                  directed = FALSE)
g.underreporting <- igraph::simplify(g.underreporting)

overreporting.df <- 
    nyakatoke[nyakatoke$willingness_link1 == 1 & nyakatoke$willingness_link2 == 1,]
g.overreporting <- igraph::graph_from_data_frame(overreporting.df, 
                                                 directed = FALSE)
missing.vertices <- c(7, 30, 32, 46, 44, 65, 84, 88, 91, 96, 107, 110, 
                      116, 117, 118, 119)
g.overreporting <- (Reduce(f = function(x, y) {y + igraph::vertex(x)},
       x = missing.vertices,
       init = g.overreporting,
       right = TRUE))
```

# Simulation parameters

For the simulations, we need to specify the size and order or average degree
(depending on the type of simulation we are performing) of the random graphs
we wish to create, as well as the number of simulations we wish to run.  The 
order of the graph will be 119 (the number of the households in the survey).  
The size of the graph will depend on the underlying graph we are trying to 
simulate (e.g. the directed/desire-to-link graph, the underreporting graph and 
the overreporting graph).

```{r simulation parameters}
sim.size <- 1000
sim.order <- 119
size.directed <- igraph::gsize(g.directed)
size.underreporting <- igraph::gsize(g.underreporting)
size.overreporting <- igraph::gsize(g.overreporting)
avg.degree.directed <- mean(igraph::degree(g.directed, mode = "out"))
avg.degree.underreporting <- mean(igraph::degree(g.underreporting))
avg.degree.overreporting <- mean(igraph::degree(g.overreporting))
```

# Simulations

We are now ready to start simulating random graphs in order to examine how their
properties line up with the observed network.  We will run three sets of
simulations:

1. The Erdős–Rényi $G(n, M)$ model, which samples uniformly from the set of all
    possible graphs of $n$ vertices and $M$ edges.  We set $n = 119$ and 
    $M$ equal to the order of the graph we are trying to simulate.
    
2. The Erdős–Rényi $G(n, p)$ model, is constructed by connecting nodes randomly.
    Each edge is included in the graph with probability $p$ independent from 
    every other edge.  We set $p$ equal to the average degree of the graph
    divided by 119.
    
3. The degree sequence graph samples uniformly from all possible graphs with
    the same degree sequence as the underlying graph we are trying to simulate.
    

## Erdős–Rényi $G(n, M)$

Here we simulate using the Erdős–Rényi $G(n, M)$ model.  We start by simulating
the underreporting model.

### Underreporting model

```{r E-R-M underreporting simulation}
underreporting.sim <- replicate(n = sim.size,
                                igraph::sample_gnm(sim.order, size.underreporting), 
                                simplify = FALSE)
```

Now let's get some basic statistics about our simulated graphs.  We start by
examining the size and order of the graph.  Since these are set as simulation
parameters, all the generated graphs should have an order of 119 and size of 490.
Also, since the size and order of the graph are fixed, the average number of 
connections per node is alo fixed at (2*490)/119 = 8.235294.

#### Size, Order and Degree

We verify this below.

```{r underreporting size order and degree}
# The order of all the graphs should be sim.order = 119 and the size should be 
# size.underreporting = 490
all(sapply(underreporting.sim, igraph::gorder) == sim.order)
all(sapply(underreporting.sim, igraph::gsize) == size.underreporting)
all(sapply(underreporting.sim, function (x) mean(igraph::degree(x)))
        ==  2*size.underreporting/sim.order)
```

#### Global Clustering Coefficient

We next examine the clustering coefficient of the generated graphs.  We start
with the "global cluserting coefficient":

$$
Cl(g) = \frac{\sum_{i; j \ne i; k \ne j, i}g_{ij} g_{jk} g_{ik}}{\sum_{i; j \ne i; k \ne j, i}g_{ij} g_{jk}}
$$

```{r underreporting clustering coefficient, fig.width=6, fig.height=5}
underreporting.cc <- sapply(underreporting.sim, 
                            function(x) igraph::transitivity(x))
summary(underreporting.cc)
round(quantile(underreporting.cc, c(0.025, 0.975)), 3)
h <- hist(underreporting.cc, plot = FALSE)
h$density <- h$counts/sum(h$counts)
plot(h, 
     freq = FALSE,
     xlab = "Clustering Coefficient",
     main = "Distribtuion of Global Clustering Coefficients
     in Erdos-Renyi simulations")
```

#### LOcal Clustering Coefficient

An alternative to the global clustering coefficient is the average local
clustering coefficient.  For each vertex on the graph, we calculate a local
clustering coefficient at the node.

$$
Cl_i(g) = \frac{\sum_{i; j \ne i; k \ne j, i}g_{ij} g_{jk} g_{ik}}{\sum_{j \ne i; k \ne j, i}g_{ij} g_{jk}}
$$

We then calculate the average clustering coefficient for the whole graph:

$$
Cl^{Avg}(g) = \frac{1}{n} \sum_i Cl_i(g)
$$

where $n$ is the order of the graph (119 in our case).

```{r underreporting local clustering, fig.width=6, fig.height=5}
underreporting.local.cc <- sapply(underreporting.sim, function(g) {
    mean(igraph::transitivity(graph = g, type = "local", isolates = "zero"))
})

round(quantile(underreporting.local.cc, c(0.025, 0.975)), 3)
h <- hist(underreporting.local.cc, plot = FALSE)
h$density <- h$counts/sum(h$counts)
plot(h, 
     freq = FALSE,
     xlab = "Local Clustering Coefficient",
     main = "Distribtuion of Local Clustering Coefficients
     in Erdos-Renyi simulations")
```

### Overreporting Model

We now repeat the analysis for the overreporting model.

```{r overreporting simulation}
overreporting.sim <- replicate(n = sim.size,
                                igraph::sample_gnm(sim.order, size.overreporting), 
                                simplify = FALSE)
```

#### Size, order and Degree

For the overreporting model, the order of all the simulated graphs should 